---
title: "Cleaning Data-Capstone"
author: "E.Matthews"
date: "14 March 2017"
output: html_document
---

```{r survey, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown- Cleaning Capstone Data

This document details the processes and code used to clean the Mental Health Survey (2014) [Data Set](www.kaggle.com/osmi/mental-health-in-tech-survey). This data set contains 1259 observations and 27 variables. This data set was obtained from the Open Source Mental Illness Project (on the Kaggle website)

Before beginning any analyses, the raw data set was reviewed to determine what cleaning/manipulating was needed. Several issues were noted including:

* *Gender variable*-- multiple misspellings, multiple forms to represent one gender (e.g. M, or m, Male for Male) as well as a number of other gender representations (Cis, femme-male, trans)

* *Wellness, Anonymity, Coworkers, Care Options, Supervisor, Coworker, Seek _Help variables* required some cleaning as the levels within the variable frequently had spaces (e.g. "Don't know")

* *State variable*-- many missing values due to responses from foreign respondents

* *Number of Employees and Age variables*-- numeric errors present (e.g. 999 for an age)

The following is the code for cleaning each variable :


#### Gender Variable

```{r warning = FALSE, error = FALSE, eval=FALSE}

read_csv("survey.csv")
b<- as.factor(survey$Gender)
levels(b)
# 45 levels of Gender-- will have them match to Male, Female, Other
survey$Gender[survey$Gender=="M"] <- "Male "
survey$Gender[survey$Gender=="m"]<-" Male "
survey$Gender[survey$Gender== "Agender"]<- "Other"
survey$Gender[survey$Gender=="All"] <- "Other"
survey$Gender[survey$Gender == "Androgyne"] <- "Other"
survey$Gender[survey$Gender == "Cis Female"] <- 'Other'
survey$Gender[survey$Gender== "cis male"] <-"Other"
survey$Gender[survey$Gender== "Cis Male"] <- "Other"
survey$Gender[survey$Gender== "Cis Man"] <- "Other"
survey$Gender[survey$Gender == "Enby"] <- "Other"
survey$Gender[survey$Gender == "f"] <-"Female"
survey$Gender[survey$Gender == "F"] <- "Female"
survey$Gender[survey$Gender== "femail"] <- "Female"
survey$Gender[survey$Gender== "Femake"] <- "Female"
survey$Gender[survey$Gender== "female"] <-"Female"
survey$Gender[survey$Gender == "Female (cis)"] <- "Other"
survey$Gender[survey$Gender== "Female (trans)"] <-"Other"
survey$Gender[survey$Gender== "fluid"] <- "Other"
survey$Gender[survey$Gender== "Genderqueer"] <- "Other"
survey$Gender[survey$Gender== "Guy (-ish) ^_^"] <- "Other"
survey$Gender[survey$Gender== "Mail"] <- "Male"
survey$Gender[survey$Gender == "maile"] <- "Male"
survey$Gender[survey$Gender == "Make"] <- "Male"
survey$Gender[survey$Gender == "male"] <- "Male"
survey$Gender[survey$Gender == "Male (CIS)"] <- "Other"
survey$Gender[survey$Gender == "male learning androgynous"]<- "Other"
survey$Gender[survey$Gender == "Male-ish"] <- "Other"
survey$Gender[survey$Gender =="Malr"] <- "Other"
survey$Gender[survey$Gender == "Man"] <- "Male"
survey$Gender[survey$Gender== "msle"] <- "Male"
survey$Gender[survey$Gender== "Nah"] <- "Other"
survey$Gender[survey$Gender== "Neuter"] <- "Other"
survey$Gender[survey$Gender == "non-binary"] <- "Other"
survey$Gender[survey$Gender== "ostensibly male, unsure what that
              really means" ] <- "Other"
survey$Gender[survey$Gender == "p"] <- "Other"
survey$Gender[survey$Gender == "queer"] <- "Other"
survey$Gender[survey$Gender =="queer/she/they"] <- "Other"
survey$Gender[survey$Gender == "something kinda male?"] <- "Other"
survey$Gender[survey$Gender == "Trans woman"] <- "Other"
survey$Gender [survey$Gender == "Trans-female"] <- "Other"
survey$Gender[survey$Gender == "woman"] <- "Female"
survey$Gender[survey$Gender == "Woman"] <- "Female"
survey$Gender[survey$Gender == "A little about you"] <- "Other"
survey$Gender[survey$Gender == "cis-female/femme"] <- "Other"
# save clean file
write.csv(survey, "survey_clean.csv")

```

#### Care_Options Variable

```{r warning=FALSE, error=FALSE, eval=FALSE}
read.csv("survey_clean.csv")
summary(survey_clean)
#Clean care options variable--there is a space between Not sure--change to unsure
survey_clean$care_options[survey_clean$care_options== "Not sure"]<- "Unsure"
survey_clean[is.na(survey_clean$care_options),]
```
#### Benefits Variable
```{r warning=FALSE, error=FALSE, eval=FALSE}
#Clean benefits--space between don't know
survey_clean$benefits[survey_clean$benefits == "Don't know"]<- "DK"

```
#### Wellness_Program , Anonymity, Coworkers, Supervisor, Mental vs Physical, Seek_Help Variables
```{r warning=FALSE, error=FALSE, eval=FALSE}
#Clean wellness program variable-- space between don'tknow
survey_clean$wellness_program[survey_clean$wellness_program == "Don't know"]<- "DK"
#Clean anonymity variable-- space between "don't know"
survey_clean$anonymity[survey_clean$anonymity == "Don't know"] <- "DK"
# Clean coworkers and supervisors columns-- some of them --spaces
survey_clean$coworkers[survey_clean$coworkers == "Some of them"] <- "Some"
survey_clean$supervisor[survey_clean$supervisor == "Some of them"] <- "Some"
#Clean mentalvs phys column--space between Don't know
survey_clean$mental_vs_physical[survey_clean$mental_vs_physical == "Don't know"]<-"DK"
#Clean seek help column-space in Don't know
survey_clean$seek_help[survey_clean$seek_help == "Don't know"]<- "DK"
```
#### Leave Variable
```{r warning=FALSE, error=FALSE, eval=FALSE}
#Clean leave column--many spaces 
survey_clean$leave[survey_clean$leave == "Don't know"]<- "DK"
survey_clean$leave[survey_clean$leave == "Very easy"] <- "Veryeasy"
survey_clean$leave[survey_clean$leave == "Somewhat difficult"]<- "Somediff"
survey_clean$leave[survey_clean$leave == "Somewhat easy"] <- "Someasy"
survey_clean$leave[survey_clean$leave == "Very difficult"]<-"Verydiff"
#save cleaned file
write_csv(survey_clean, "survey_clean1.csv")
```

#### Age Variable-- multiple errors in Age column (age outliers such as 5 or 99999) . These were converted to NA, and then replaced with the mean age

```{r warning=FALSE, error= FALSE, eval =FALSE}
survey_clean2 <-read.csv("survey_clean1.csv")
View(survey_clean2)
# Fix age variable
a<-as.factor(survey_clean2$Age)
levels(a)
survey_clean2$Age[(survey_clean2$Age == -1726) |
                    (survey_clean2$Age == -1) |
                    (survey_clean2$Age == 5) |
                    (survey_clean2$Age == 329) |
                    (survey_clean2$Age == 99999999999)] <-NA
survey_clean2$Age[(survey_clean2$Age == -29)] <- 29
View(survey_clean2)
mean(survey_clean2$Age)
# mean age is 32- replace NA with mean age 
survey_clean2$Age[is.na(survey_clean2$Age)] <- mean(survey_clean2$Age, na.rm= TRUE)

```


#### Clean Number of Employees Column 
```{r warning = FALSE, error= FALSE, eval= FALSE}
# Fix Number of Employees Column
levels(survey_clean2$no_employees)
survey_clean2$no_employees <- as.character(survey_clean2$no_employees)
survey_clean2$no_employees[(survey_clean2$no_employees == "Jun-25")]<-"6-25"
survey_clean2$no_employees[(survey_clean2$no_employees == "01-May")]<- "1-5"
write.csv(survey_clean2, "survey_clean2.csv")
```
#### Change NA to Foreign in State Column

```{r warning=FALSE, error= FALSE, eval= FALSE}
class(survey_clean4$state)
levels(survey_clean4$state)
class(survey_clean4$Country)
levels(survey_clean4$Country)
# change state NA to foreign
survey_clean4$state <-as.character(survey_clean4$state)
class(survey_clean4$state)
survey_clean4$state[is.na(survey_clean4$state)]<-"Foreign"

write_csv(survey_clean4, "survey_clean4.csv")
```

#### Add new column for Region based on State info

```{r warning=FALSE, error= FALSE, eval= FALSE}
survey_clean4 <- read.csv("survey_clean4.csv")
# add new column - region (for states)
survey_clean4$region <- ifelse(survey_clean4$state == "NY", "NE",
ifelse(survey_clean4$state == "PA","NE",
ifelse(survey_clean4$state == "RI","NE",                            ifelse(survey_clean4$state == "CT", "NE",
                                                    ifelse(survey_clean4$state == "NJ", "NE",
                                                           ifelse(survey_clean4$state == "NH", "NE",
                                                                  ifelse(survey_clean4$state == "MA", "NE",
                                                                    ifelse(survey_clean4$state == "ME", "NE",
                                                                    ifelse(survey_clean4$state == "VT", "NE",
                                                                    ifelse(survey_clean4$state == "MN", "MW",
                                                                    ifelse(survey_clean4$state == "MI", "MW",
                                                                     ifelse(survey_clean4$state == "MO", "MW",
                                                                      ifelse(survey_clean4$state == "KS", "MW",
                                                                                                                   ifelse(survey_clean4$state =="NE", "MW",
                                                                                                                          ifelse(survey_clean4$state == "IN", "MW",
                                                                                                                                 ifelse(survey_clean4$state == "IL", "MW",
                                                                                                                                        ifelse(survey_clean4$state == "SD", "MW",
                                                                                                                                          ifelse(survey_clean4$state == "OH", "MW",
                                                                                                                                          ifelse(survey_clean4$state == "IA","MW",
                                                                                                                                          ifelse(survey_clean4$state == "WI", "MW",
                                                                                                                                        ifelse(survey_clean4$state == "FL", "S",
                                                                                                                                                                        
                                                                                                                                        ifelse(survey_clean4$state == "AL", "S",
                                                                                                                                        ifelse(survey_clean4$state == "GA", "S",
                                                                                                                                                                                         ifelse(survey_clean4$state == "SC", "S",
                                                                                                                                                                                                ifelse(survey_clean4$state == "TN", "S",
                                                                                                                                                                                                       ifelse(survey_clean4$state == "TX", "S",
                                                                                                                                                                                                              ifelse(survey_clean4$state == "OK", "S",
                                                                                                                                                                                                              ifelse(survey_clean4$state == "LA", "S",
                                                                                                                                                                                                              ifelse(survey_clean4$state == "DC", "S",
                                                                                                                                                                                                                ifelse(survey_clean4$state == "KY", "S",
                                                                                                                                                                                                                ifelse(survey_clean4$state == "MD", "S",
                                                                                                                                                                                                                ifelse(survey_clean4$state == "NC", "S",
                                                                                                                                                                                                               ifelse(survey_clean4$state == "WV", "S",
                                                                                                                                                                                                                                                               ifelse(survey_clean4$state == "VA", "S",
                                                                                                                                                                                                                                                                      ifelse(survey_clean4$state == "MS", "S",
                                                                                                                                                                                                                                                                             ifelse(survey_clean4$state == "ID", "W",
                                                                                                                                                                                                                                                                                    ifelse(survey_clean4$state == "AZ", "W",
                                                                                                                                                                                                                                                                                    ifelse(survey_clean4$state == "NM", "W",
                                                                                                                                                                                                                                                                                    ifelse(survey_clean4$state == "WY", "W",
                                                                                                                                                                                                                                                                                    ifelse(survey_clean4$state == "CO", "W",
                                                                                                                                                                                                                                                                                    ifelse(survey_clean4$state == "UT", "W",
                                                                                                                                                                                                                                                                                    ifelse(survey_clean4$state == "NV", "W",
                                                                                                                                                                                                                                                                                    ifelse(survey_clean4$state == "CA","W",
                                                                                                                                                                                                                                                                                                                                     ifelse(survey_clean4$state =="OR", "W", "foreign"
                                                                                                                                                                                                                                                                                                                                            ))))))))))))))))))))))))))))))))))))))))))))
# reorder columns for clarity
survey_clean5 <- survey_clean4[,c(1:7, 30, 8:29)]
View(survey_clean5)
write.csv(survey_clean5, "survey_clean5.csv")
```
